### 2018-09-21 manuscript/methods.tex
thinking about model names
edited model descriptions

We first defined a binary state speciation and extinction model (BiSSE, \citet{maddison_2007})  for polyploidy evolution where taxa was classified as diploid (D=0) and polyploid (P=1) \cref{table:stateclassifications}.
In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$) , net diversification rates ($r_D=\lambda_D-\mu_D, r_P=\lambda_P-\mu_P$), and  relative extinction rates ($\nu_D=\frac{\mu_D}{\lambda_D}, \nu_D=\frac{\mu_D}{\lambda_D}$) just as previously explored in \citet{mayrose_2011}.
The cladogenetic changes explored in BiSSE were polyploidization rate are represented by parameter $\rho$ and by diploidization parameter rate $\delta$.
% E: cladogenetic?

% E: above and below, drop numerical state names ... even in the table

As a second step, we fitted a hidden state speciation and extinction model (HiSSE, \citet{beaulieu_2016}) to evaluate whether the differences in diversification rates were found due to a hidden trait associated to polyploidy.
As discussed by \citet{beaulieu_2016} BiSSE-like models suffer from a large type I error because they fail to account as part of the null hypothesis heterogeneous diversification rate changes  that do not depend on the trait of interest.
By including a hidden trait in a state speciation and extinction  model,  HiSSE-like models address the heterogeneous background diversification changes while also parsing out the possible signal of diversification due to the trait of interest. 
Therefore, HiSSE model for polyploidy linked to diversification has four states: diploid and polyploid subdivided by a binary hidden trait with states A and B for which we estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\lambda_{D_B}, \lambda_{P_A},\lambda_{P_B}$), extinction rates ($\mu_{D_A},\mu_{D_B}, \mu_{P_A},\mu_{P_B}$),  net diversification rates ($r_{D_A},r_{D_B},r_{P_A},r_{P_B}$), and relative extinction($\nu_{D_A},\nu_{D_B}, \nu_{PA},\nu_{P_B}$).
The cladogenetic changes assumed by the fitted model are polyploidization rate $\rho$, diploidization rate $\rho$, and we assumed that the changes between every hidden state are symmetrical with rate $\alpha$.

For breeding system both BiSSE and a HiSSE models were fitted independently by coding data as self-compatible (SC=0) or self-incompatible (SI=1) \cref{table:stateclassifications}.
The BiSSE model fitted to breeding system data was done with the goal of investigating if the effect of self-incompatibility in diversification was similar to the pattern found by \citet{goldberg_2012}.
However, the HiSSE model on breeding system parses out if the effect of breeding system in diversification is significant  by minimizing type I error as mentioned above.
For these models we assumed that self-compatibility is irreversible as discussed by \citet{igic_2013}, so we defined the transition rate $q_{IC}$ as the parameter for changes from self-incompatible to self-compatible state.
% E: explain irreversibility assumption; fix ancestry instead?

Next, we proposed a multivariate speciation and extinction model (MuSSE, \citet{fitzjohn_2012}) to investigate the link amongst diversification, breeding system, and polyploidy simultaneously.
The MuSSE model is defined using three states self-compatible diploids (SC-D=0), polyploids that are always self-compatible (ref for this??) (P=1), and self-incompatible diploids (SI-D=2, see table \cref{table:stateclassifications}) and contains ten parameters, six defining diversification ($\lambda_{SD}, \mu_{SD},\lambda_{P},\mu_{P}, \lambda_{ID},\mu_{ID}$) and the other four key cladogenetic changes that are: polyploidization of self-compatible diploids $\rho_{SD}$, diploidization $\delta$, polyploidization of self-incompatible diploids $\rho_{ID}$, and self-incompatible to self-compatible rate $q_{IC}$.
% E: for SC, use C instead of S
% E: lack of SI-P should maybe be explained in Intro?

Since the null hypothesis of the MuSSE model is that the diversification is equal and constant for all three states defined, it is possible that MuSSE can also suffer from large type I errors.
In order to account for heterogeneity of the diversification rates and parse the signal in diversification coming from breeding system and polyploidy we extended MuSSE model to account for a hidden state, as \citet{beaulieu_2016} did in the bivariate case.
The model we proposed is a multivariate and hidden states speciation and extinction stochastic process (MuHiSSE) that allowed us to account simultaneously for the diversification rates linked to breeding system and polyploidy but also the presence of some more heterogeneity in the process.
This model is analogous to the GeoHiSSE model proposed by Caetano et al. (2018) in a biogeographical context.
The full MuHiSSE model has 26 parameters, however, our goal was to look for diversification rate differences so we fitted a simplified version of 17 parameters by fixing the rates amongst hidden states to be equal with parameter $\alpha$ and the transition rates amongst breeding system and polyploidy as defined in the MuSSE model  ($\rho_{SD}$,  $\delta$,  $\rho_{ID}$,  $q_{IC}$) despite the hidden state.
Using MuHiSSE we estimated twelve speciation and extinction rates ($\lambda_{SD_A}, \mu_{SD_A},\lambda_{P_A},\mu_{P_A}, \lambda_{ID_A},\mu_{ID_B},\lambda_{SD_B}, \mu_{SD_B},\lambda_{P_B},\mu_{P_B}, \lambda_{ID_B},\mu_{ID_B}$), and the net diversification and relative extinction rates associated with them.
% E: caetano ref; and huang

>>>>

% E: I think it would help to start with an "overall strategy" paragraph, maybe something like this.
To investigate the associations of these traits with lineage diversification, we fit six models from the BiSSE \citep{maddison_2007} family.
The simplest models consider ploidy only or breeding system only, as previous work has done. % add refs
We also consider the two traits simultaneously, to investigate whether the effect of one might override the effect of the other.
Additionally, to allow for the possibility that something other than these traits might affect diversification, we add models that include a hidden trait. % hisse ref
% E: And say something about including diploidization?  Will the motivation be in the Intro?
% E: And explain about the SC -> SI rate being fixed to 0, so we don't bury the explanation below? vs fixing ancestry?

We first considered only the ploidy trait, scoring each species in either the diploid state (D) or the polyploid state (P).
To allow for the influence of this binary trait on lineage diversification, we used the BiSSE model \citep{maddison_2007}.
% \cref{table:stateclassifications}
In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$), net diversification rates ($r_D=\lambda_D-\mu_D,\ r_P=\lambda_P-\mu_P$), and relative extinction rates ($\nu_D = \mu_D / \lambda_D,\ \nu_D = \mu_D / \lambda_D$) for each state.
% as previously explored by \citet{mayrose_2011} % E: actually, they used bisse-ness/classe to allow for cladogenetic change
We also simultaneously estimated the polyploidization rate (the transition rate from D to P, $\rho$), and we allowed for diploidization (the transition rate from P to D, $\delta$).
We call this the ``ploidy only'' model. % E: not sure that we need to be this explicit, but try for now

Second, we considered ploidy and also a unobserved, hidden trait.
As discussed by \citet{beaulieu_2016}, the previous model might attribute diversification differences to the ploidy trait merely because there are no other explanatory factors in the model.
% As a second step, we fitted a hidden state speciation and extinction model (HiSSE, \citet{beaulieu_2016}) to evaluate whether the differences in diversification rates were found due to a hidden trait associated to polyploidy.
% As discussed by \citet{beaulieu_2016} BiSSE-like models suffer from a large type I error because they fail to account as part of the null hypothesis heterogeneous diversification rate changes  that do not depend on the trait of interest.
By including a hidden trait in a state speciation and extinction model, HiSSE-like models allow for heterogeneous background diversification while also parsing out the possible signal of diversification due to the trait of interest. 
Our second model thus has four states, with D and P subdivided by an unobserved binary trait with states A and B.
We therefore estimated the posterior probability distributions of four speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\  \lambda_{P_A},\ \lambda_{P_B}$), four extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), and the corresponding four net diversification rates and relative extinction rates.
We again include transition due to polyploidization (rate $\rho$) and diploidization (rate $\delta$), and we assume that the changes between hidden states are symmetrical with rate $\alpha$.
We call this the ``ploidy and hidden trait'' model.

% E: I added the names of the rate parameters to these next models

In our third, ``breeding system only'' model, each species is scored as either self-incompatible (SI) or self-compatible (SC).
In this BiSSE analysis, state-dependent speciation rates ($\lambda_I$, $\lambda_C$) and extinction rates ($\mu_I$, $\mu_C$) are estimated, with the corresponding net diversification and relative extinction rates.
The rate for transitions from SI to SC is $q_{IC}$, and regain of SI is prohibited, as explained above.
This is the same as the analysis of \citet{goldberg_2010}.

The fourth model considers both breeding system and a hidden trait, again using a HiSSE strategy to test whether diversification shifts unrelated to the focal trait might be attributed to it by the previous BiSSE analysis.
In this ``breeding system and hidden trait'' model, there are four speciation rates ($\lambda_{I_A},\ \lambda_{I_B},\  \lambda_{C_A},\ \lambda_{C_B}$), four extinction rates ($\mu_{I_A},\ \mu_{I_B},\ \mu_{C_A},\ \mu_{C_B}$), and correspondingly four rates of net diversification and relative extinction.
Transition rates are again $q_{IC}$ for loss of SI and $\alpha$ for changes in the hidden state.

% E: I adjusted the rate names a little.  Does this seem okay?
Our fifth model investigates the link among diversification, breeding system, and ploidy simultaneously.
There are three possible states: self-incompatible diploids (ID), self-compatible diploids (CD), and self-compatible polyploids (CP); self-incompatible polyploids are not present in this family, as explained earlier.
This ``ploidy and breeding system'' MuSSE model \citep{fitzjohn_2012} has six rates for speciation and extinction ($\lambda_{ID},\ \mu_{ID},\ \lambda_{CD},\ \mu_{CD},\ \lambda_{CP},\ \mu_{CP}$).
It also has four rates for transitions:
polyploidization of SI diploids (transition from ID to CP at rate $\rho_I$),
polyploidization of SC diploids (transition from CD to CP at rate $\rho_C$),
diploidization, which does not restore SI (transition from CP to CD at rate $\delta_C$),
and loss of SI without a change in ploidy (transition from ID to CD at rate $q_{IC}$).

Finally, the sixth model considers a hidden trait along with ploidy and breeding system.
This again provides a means to account simultaneously for diversification rates linked to ploidy and breeding system while also allowing for the presence of heterogeneity from unknown sources.
(A similar approach was used for a different variant of BiSSE by \citealt{huang_2018,caetano_2018}.)
The full model would have 26 parameters, but our goal is to look for diversification rate differences so we fit a simplified version with 17 parameters by fixing the transition rates among hidden states to be equal with parameter $\alpha$, and the transition rates between breeding system and ploidy states as defined in the previous model ($\rho_I$, $\rho_C$, $\delta_C$,  $q_{IC}$) independent of the hidden state.
We estimated twelve speciation and extinction rates ($\lambda_{ID_A},\ \mu_{ID_A},\ \lambda_{CD_A},\ \mu_{CD_A},\ \lambda_{CP_A},\ \mu_{CP_A},\ \lambda_{ID_B},\ \mu_{ID_B},\ \lambda_{CD_B},\ \mu_{CD_B}, \lambda_{CP_B},\ \mu_{CP_B}$), and the net diversification and relative extinction rates associated with them.
